### 2017-11-12 manuscript.rmd
refined model discussion, did not update pdf

We did not find increased evolutionary change in expression following duplication events as compared to speciation events (Fig. 2*B*). The Wilcoxon rank test does not reject the null hypothesis that the rate of evolution following duplications is the same as or less than the rate following speciation (p value = `r wilcox_test_result`). Our phylogenetic comparative analysis, unlike the previously published pairwise comparative analysis [@KryuchkovaMostacci:2016iw], therefore finds no support for the ortholog conjecture in this system.

@@

Levin *et al.* [@Levin:2016bp] arrived at this conclusion by making multiple pairwise comparisons of ortholog expression data sampled throughout the course of embryonic development. For each species pair, they identified the orthologs shared by these species. This list of shared genes was different from species pair to species pair. They characterized each of these orthologs in each species as having expression that peaks in early, mid, or late temporal phase of development. They then calculated a similarity score for each temporal phase for each species pair based on the fraction of genes that exhibited the same patterns in each species. The distributions of similarity scores are plotted in their Fig. 4d [@Levin:2016bp], and their Kolmogorov–Smirnov (KS) tests indicated that the early distribution and late distribution were each significantly different from mid distribution (P < $10^{-6}$ and P < $10^{-12}$, respectively). This is the support they presented for the inverse hourglass model.

@@

Thanks to Steve Haddock, Alex Damian Serrano, August Guang, Bruno Vellutini, Zack Lewis, and Matthew Hahn for feedback on the manuscript. Marc Robinson-Rechavi provided very helpful information about the KMRR analyses and valuable suggestions for our manuscript. The Ensembl Comparative Genomics team, including Matthieu Muffato, provided excellent support for helping us better understand and use their Compara trees. CWD's contribution was supported by the National Science Foundation (DEB-1256695 and the Waterman Award). AH was supported by the European Research Council Community’s Framework Program Horizon 2020 (2014–2020) ERC grant agreement 648861.
\pagebreak

@@

#### Model selection and sensitivity

We removed the `r k_percentile` of genes with the lowest phylogenetic signal, corresponding to a K cutoff for retained genes of greater than `r signif( k_thresh, 3)`. Repeating the phylogenetic analyses on this subset of genes with strongest phylogenetic signal for Tau still do not reject the null hypothesis (Wilcoxon p = $`r  wilcox_test_result_k %>% round( 3 )`$).

@@

>>>>

We did not find increased evolutionary change in expression following duplication events as compared to speciation events (Fig. 2*B*). The Wilcoxon rank test does not reject the null hypothesis that the rate of evolution following duplications is the same as or less than the rate following speciation (p value = `r wilcox_test_result`). Our phylogenetic comparative analysis, unlike the previously published pairwise comparative analysis [@KryuchkovaMostacci:2016iw], therefore finds no support for the ortholog conjecture in this system. This result holds when we exclude the genes with the lowest phylogenetic signal according to Blomberg's $K$ [@Blomberg:2003jg] and is robust to model selection (*SI Appendix*).

@@

Levin *et al.* [@Levin:2016bp] arrived at this conclusion by making multiple pairwise comparisons of ortholog expression data sampled throughout the course of embryonic development. For each species pair, they identified the orthologs shared by these species. This list of shared genes was different from species pair to species pair. They characterized each of these orthologs in each species as having expression that peaks in early, mid, or late temporal phase of development. They then calculated a similarity score for each temporal phase for each species pair based on the fraction of genes that exhibited the same patterns in each species. The distributions of similarity scores are plotted in their Fig. 4d [@Levin:2016bp], and their Kolmogorov-Smirnov (KS) tests indicated that the early distribution and late distribution were each significantly different from mid distribution (P < $10^{-6}$ and P < $10^{-12}$, respectively). This is the support they presented for the inverse hourglass model.

@@

Thanks to Steve Haddock, Alex Damian Serrano, August Guang, Bruno Vellutini, Zack Lewis, and Matthew Hahn for feedback on the manuscript. Marc Robinson-Rechavi provided very helpful information about the KMRR analyses and valuable suggestions for our manuscript. The Ensembl Comparative Genomics team, including Matthieu Muffato, provided excellent support for helping us better understand and use their Compara trees. CWD's contribution was supported by the National Science Foundation (DEB-1256695 and the Waterman Award). AH was supported by the European Research Council Community's Framework Program Horizon 2020 (2014-2020) ERC grant agreement 648861.
\pagebreak

@@

#### Phylogenetic signal and models

```{r Fig_SXX, fig.width=5, fig.height=3, fig.cap="Density plot of Blomberg's K. The shaded portion of the plot indicates the gene trees that were analyzed after excluding the trees that fell below the K threshold."}

	max_k = 2
	dens = density(tree_summary$K[ !is.na(tree_summary$K) ], from = 0, to = max_k)
	D = data.frame(x = dens$x, y = dens$y)
	thresh = k_thresh
	shade = rbind(c(max_k, 0), c(thresh,0), subset(D, x > thresh))

	thresh_label = paste( "\nthreshold: > ", round(thresh,2), sep='' )

	g_k_density = 
		ggplot(D, aes(x,y)) + 
			geom_line() + 
			geom_polygon( data = shade, aes(x, y) ) +
			xlab("K") +
			ylab("density") +
			theme_classic() +
			geom_vline(xintercept = thresh, col="grey") +
			geom_text(aes(x=thresh, label=thresh_label, y=mean(range(y))), colour="grey", angle=90) +
			ggtitle( "A Density of K" )

	g_k_density

The gene trees considered here vary widely in their phylogenetic signal for Tau according to Blomberg's $K$ [@Blomberg:2003jg] (*SI Appendix*, Fig. XX). $K$ describes the tendancy for closely related tips to resemble each other, where $K=0$ indicates that close relatives resemble each other no more than they do distant relatives and $K=1$ indicates that the data have strong phylogenetic signal. To rule out the possibility that gene trees with little phylogenetic signal were diluting support for the Ortholog Conjecture, we removed the `r k_percentile` of genes with the lowest phylogenetic signal, corresponding to a $K$ cutoff for retained genes of greater than `r signif( k_thresh, 3)`. Repeating the phylogenetic analyses on this subset of genes with strongest phylogenetic signal for Tau still do not reject the null hypothesis (Wilcoxon p = $`r  wilcox_test_result_k %>% round( 3 )`$). This indicates that our inability to reject the null hypothesis with phylogenetic methods is not due to a lack of phylogenetic signal.

Phylogenetic independent contrasts (PICs) are based on a Brownian Motion (BM) model of character evolution [@Felsenstein:1985ua]. PICs could fail to reject the null hypothesis if BM is not a good model for the evolution of Tau on the gene phylogenies. We therefore compared the fit of BM to the more complex Ornstein Uhlenbeck (OU) model of evolution.

@@

These analyses indicate that our inability to reject the null hypothesis is robust under different models of Tau evolution.
